---
title: Introduction to distance sampling
author: Centre for Research into Ecological and Environmental Modelling  **University of St Andrews**
subtitle: Workshop, 04-15 May 2020
date: Exercise 5. Point transect sampling 
output: 
  html_document:
    number_sections: yes
    theme: flatly
    highlight: tango
    df_print: kable
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

<div class="alert  alert-success">
  <strong>Solutions</strong> Point transect exercises
</div>

# Simulated data

The code for importing and checking these data and fitting various models is shown below. 

```{r, fig.width=4, fig.height=4, message=FALSE}
library(Distance)
# Read in data
data("PTExercise")
# What does data look like
head(PTExercise, n=3)
conversion.factor <- convert_units("meter", NULL, "hectare")
# Fit half-normal detection function, no truncation
PTExercise.hn <- ds(data=PTExercise, transect="point", key="hn",
               convert.units=conversion.factor)
plot(PTExercise.hn, pdf=TRUE)

# Try truncation of 20m based on preliminary fit

# Half normal, no adjustments
PTExercise.hn.t20m <- ds(data=PTExercise, transect="point", key="hn", truncation=20,
                    convert.units=conversion.factor)
# Hazard rate, no adjustments
PTExercise.hr.t20m <- ds(data=PTExercise, transect="point", key="hr", truncation=20,
                    convert.units=conversion.factor)
# Uniform, cosine adjustments
PTExercise.uf.cos.t20m <- ds(data=PTExercise, transect="point", key="unif", 
                        adjustment="cos", truncation=20,convert.units=conversion.factor)
```


```{r, echo=F}
# Same caveat as with previous exercises, do not get excited
#   about the code in this chunk; it is not necessary for your 
#   understanding of distance sampling.
pt.tab <- data.frame(DetectionFunction=c("Half-normal","Half-normal",
                                         "Hazard rate","Uniform"), 
                     Adjustments=c("None","None","None","Cosine"), Truncation=c(34.2,20,20,20), 
                     AIC=rep(NA,4), Density=rep(NA,4), D.CV=rep(NA,4), Lower.CI=rep(NA,4), Upper.CI=rep(NA,4))

get.results.f <- function(fit.model) {
  return(c(AIC=summary(fit.model$ddf)$aic,
         D=fit.model$dht$individuals$D$Estimate,
         D.CV=fit.model$dht$individuals$D$cv,
         lCL=fit.model$dht$individuals$D$lcl,
         uCL=fit.model$dht$individuals$D$ucl))
}

pt.tab[1,4:8] <- get.results.f(PTExercise.hn)
pt.tab[2,4:8] <- get.results.f(PTExercise.hn.t20m)
pt.tab[3,4:8] <- get.results.f(PTExercise.hr.t20m)
pt.tab[4,4:8] <- get.results.f(PTExercise.uf.cos.t20m)

# Print results
knitr::kable(pt.tab, caption="Results from simulated point transect data.", digits=3)
```

```{r, echo=FALSE, eval=FALSE}
# Plot detection functions
par(mfrow=c(2,2))
plot(PTExercise.hn, main="Half normal, no truncation")
plot(PTExercise.hn.t20m, main="Half normal, truncation 20m")
plot(PTExercise.hr.t20m, main="Hazard rate, truncation 20m")
plot(PTExercise.uf.cos.t20m, main="Uniform with cosine, truncation 20m")
```

```{r}
# Plot probability density functions
par(mfrow=c(2,2))
plot(PTExercise.hn, main="Half normal, no truncation", pdf=TRUE)
plot(PTExercise.hn.t20m, main="Half normal, truncation 20m", pdf=TRUE)
plot(PTExercise.hr.t20m, main="Hazard rate, truncation 20m", pdf=TRUE)
plot(PTExercise.uf.cos.t20m, main="Uniform with cosine, truncation 20m", pdf=TRUE)
```

We see a fair degree of variability between analyses - reliable analysis of point transect data is more difficult than for line transect data. We see greater loss in precision from truncating data relative to line transect sampling, but if we don't truncate data, different models can give widely differing estimates.

# Wren data (Optional)

In the code provided below, each dataset is loaded and then detection functions that we selected are fitted. 

```{r, echo=T, eval=T}
# Wren data for Method 1 (5 min counts)
data("wren_5min")
# Wren data for Method 2 (snapshot)
data("wren_snapshot")
conversion.factor <- convert_units("meter", NULL, "hectare")
# Method 1
wren5min.uf.cos.t110 <- ds(data=wren_5min, key="unif", adjustment="cos", 
                        transect="point", truncation=110, 
                        convert.units=conversion.factor)
# Method 2
wrensnap.hr.cos.t110 <- ds(data=wren_snapshot, key="hr", adjustment=NULL, 
                        transect="point", truncation=110, 
                        convert.units=conversion.factor)
```

```{r, echo=F}
# Harvest results
n <- 2
wren.tab <- data.frame(Method=1:n, Density=rep(NA,n), 
                       Lower.CI=rep(NA,n), Upper.CI=rep(NA,n))

get.results.f <- function(fit.model) { return(c(D=fit.model$dht$individuals$D$Estimate,
         lCL=fit.model$dht$individuals$D$lcl,
         uCL=fit.model$dht$individuals$D$ucl))
}

wren.tab[1,2:4] <- get.results.f(wren5min.uf.cos.t110)
wren.tab[2,2:4] <- get.results.f(wrensnap.hr.cos.t110)

# Print results
knitr::kable(wren.tab, caption="Winter wren density estimates from 5 minute counts and snapshot moment.", digits=3)
```

```{r, fig.height=4}
# Plot detection functions
par(mfrow=c(1,2))
plot(wren5min.uf.cos.t110, main="5 minute count")
plot(wrensnap.hr.cos.t110, main="Snapshot moment")
```

As the detection distance histograms indicate, winter wren showed evidence of observer avoidance, more than other species in the Montrave study.  We show the detection function graph rather than the PDF to emphasise the evasive movement aspect of the data.  If you conduct the goodness of fit test, using `gof_ds()`, you will find that the models still suitably fit the data.


